---
title: "Class 14: Genome Informatics and High Throughput Sequencing"
author: Paige Dubelko
date: February 21, 2019
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Asthma SNPs in MXL population
Read in the csv file we downloaded from [Ensembl](http://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=362656008#373531_tablePanel)
```{r}
mxl <- read.csv("class14MXL.csv", row.names = 1)
head(mxl)
```

How many G|G genotypes are in this set? Perecent values?
```{r}
table(mxl$Genotype..forward.strand.)

#table(mxl$Genotype..forward.strand.)/nrow(mxl)*100

```

## RNA-Seq

Fastq quality scores

Q. Does the first sequence have good quality?
```{r}
#install.packages("seqinr")
#install.packages("gtools")

q <- "DDDDCDEDCDDDDBBDDDCC@"
library(seqinr)
library(gtools)
asc(s2c(q)) - 33 #bc of offset
```

## Population scale analysis

Does the genotype for rs9067378 influence expression of the important asthema linked genes we found?

```{r}
geno <- read.table("230Samples.txt")
head(geno)
```

Find how many patients are G|G?
```{r}
table(geno$geno)
```

Find expression values for G|G
```{r}
inds <- geno$geno == "G/G"
summary(geno$exp[inds])
```
Find expression values for A|A
```{r}
inds.aa <- geno$geno == "A/A"
summary(geno$exp[inds.aa])
```
Find expression values for A|G
```{r}
inds.ag <- geno$geno == "A/G"
summary(geno$exp[inds.ag])
```

Create a figure that shows the distribution of these expression value
```{r}
#want expression spit by genotype
graph <- boxplot(exp ~ geno, geno)
```



