---
title: "Class 6"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Section 1: Reading files 
We are using the **read.table()** funtion and friends to read some example flat files. 

```{r}
read.csv("test1.txt")

```

For the test1 file, we are able to use read.table(), but because we notice it is separated by "," and appears to have a header, the **read.csv** is a simpler option.

We will now look at our **second** text file...
We see that this file is separated by '$' and has a header.
```{r}
read.table("test2.txt", header = TRUE, sep = "$")
```
Because we do not have a friend function that separates by "$" it is easy to edit table to specify it for this file.

Now onto the **third** file. We notice that the data is separated by what looks like spaces and does not have a header. So we just use the **read.table()** function.
```{r}
read.table("test3.txt")
```

### Section 2: R Functions
Lets write a simple function, add()...
```{r}
add <- function(x,y=1){
  #Sum the input x and y
  x + y
}
```
Now lets try add()!
```{r}
x <- 7
y <- 3
add(7,3)
```
We are also able to use our add function with vectors :) Remember: the y variable is optional.
```{r}
add(c(1,2,3))
```

If you find yourself doing the same thing three or more times, it might be time to think about writing a function. 

**2nd Function: Rescale**
```{r}
rescale <- function(x){
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}
```
Lets test on a small example, where we know what the answer should be
```{r}
rescale(1:10)
```

With how it is currently written, our rescale function will not take NA values, so we must change it.
```{r}
rescale2 <- function(x){
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
```
This will now work for vectors containing NA values because we set **na.rm** equal to TRUE
```{r}
x <- c(1,2,NA,3,10)
rescale2(x)
```
Lets edit this function one more time.
```{r}
rescale3 <- function(x, na.rm = TRUE, plot = FALSE){
  rng <- range(x, na.rm = na.rm)
  print("Hello")
  
  answer <- (x - rng[1]) / (rng[2] - rng[1])
  
  print("is it me you are looking for?")
  
  if(plot){
    plot(answer, typ = "b", lwd = 4)
  }
  print("I can see it in ...")
  return(answer)
}
```
Lets try it!
```{r}
rescale3(c(1:6), plot = TRUE)
```

